\pagestyle{empty}% Set page style to empty
%\pdfbookmark[0]{Acknowledgements}{Acknowledgements} % Bookmark name visible in a PDF viewer
\addcontentsline{toc}{chapter}{\numberline{}Acknowledgements}%

\bigskip

\begin{center}
	\Huge \textsc{\textbf{Acknowledgements}}
\end{center}

This thesis would not have been possible without the guidance and good company by a number of wonderful people. First and foremost, I would like to thank my main PhD supervisor \textit{J\"{o}rn} for taking me up as a student, giving me the freedom to follow my own ideas and for your guidance on being an academic. I hold in dear memory the dinner evenings at your house and our pub hikes through the South Downs national park. To this day I still wonder how and when you got past me at the Beachyhead Marathon. Certainly there will be a time and place for a rerun (maybe not over 26.2 miles). I will always be proud to be Scharlemann lab PhD graduate \texttt{\#2}. Thanks J\"{o}rn.
I would also like to thank my second PhD supervisor \textit{Pedram} for being there for me and giving me advice whenever I needed it. I remember the words of wisdom you told me at the start of my PhD journey, that "obtaining a PhD is not measured by the amount of accomplishments achieved or manuscripts published, but by learning how to think". I hope I have succeed in this endeavour to some degree. Looking forward meeting you, the newest addition to the Rowhani family and Lord Rusby again in the near future. 

Doing a PhD \textendash\ and especially one that is fully computational \textendash\ is often an exhausting experience. I want to express my gratitude to a number of individuals at the evolution, behaviour and environment (EBE) section at Sussex that provided me with \textendash\ the sometimes desperately \textendash\ needed (non-)academic distraction. Thanks \textit{Dan} for always being there when I needed to talk to someone. I am going to miss our chats about academic life, the nights out and cozy pub retreats. Also totally not being jealous of your postdoc fieldwork sites. Looking forward catching up with you in the future. A very big thank you goes to \textit{Owen} for his company and friendship in the office, pub, on nights out and at conferences during the second half of my PhD. Thanks for being the best office buddy a PhD student can think of. Looking forward to your visit in Vienna (but please practise your German).

I would also like to thank my other office mates during my time at Sussex. Thanks \textit{Rich} for being you and having those funny moments and distractions with me during the week. Thank you \textit{Edwin} for all the discussions and for introducing me to Bolivian-style Chili sin Carne. A special thank you also goes to the guppy lab, specifically \textit{Mijke}, \textit{Jim} and \textit{Josie}. I am going to miss the regular friday office songs, the hourse party fun, runches, super smash brothers (Jim, we still have a score to settle) and the excitement of \textendash\ almost \textendash\ winning the University of Sussex volleyball tournament. Me and Owen missed you dearly since you guys left.

A big thank you also goes to all the other nice people I had the pleasure of meeting at Sussex, such as \textit{Adam}, \textit{Jeremy N.}, \textit{Chris S. \& Lene}, \textit{Claudia \& Benj}, \textit{Jenny}, \textit{Craig}, \textit{Beth N.}, \textit{Beth G.}, \textit{Gigi} \& \textit{Toby}, \textit{Veronica}, \textit{George}, \textit{Will}, \textit{Tom}, \textit{Paul D.}, \textit{Tanya} \& \textit{Chris A.}, the \textit{Mammal society crew} and of course \textit{Rasmus} as well as many many more that I cannot possibly list here. I am certain I will see many of you again someday. Something that is often forgotten is the amount of work of the many individuals behind the scenes. This thesis would not been possible without the millions of observational biodiversity records of species and individuals collated by (non-)academics across the world in the PREDICTS and BBS projects. The same is true for the Earth observation data provided by the National Aeronautics and Space Administration (NASA) and I continue to be excited about being part of the thriving and innovative global remote sensing community. I would like to end this section by also specifically expressing my gratitude to the kitchen personnel of the Swan Inn, Falmer for providing me with that steady influx of these delicious golden delights and amber coloured beverages throughout my PhD. 

Zuletzt w\"{u}rde ich gerne meiner Familie danken für all die Jahre an Unterstützung und Rat ohne welchen ich nicht da w\"{a}re wo ich heute bin. Danke \textit{Mutsch} f\"{u}r dein offenes Ohr und deinen Auftrag an mich die Welt zu verbessern und bestehendes Wissen zu hinterfragen. Danke \textit{Vatsch} und \textit{JoJo} f\"{u}r die netten Stunden und das Gef\"{u}hl von "daheim sein". Freue mich schon auf die n\"{a}chsten Familientreffen in Aachen, Wien oder wo auch immer. 