\pagestyle{empty}% Set page style to empty
%\pdfbookmark[0]{Acknowledgements}{Acknowledgements} % Bookmark name visible in a PDF viewer
\addcontentsline{toc}{chapter}{\numberline{}Acknowledgements}%

\begin{center}
	\Huge \textsc{\textbf{Acknowledgements}}
	\hrulefill
\end{center}

First of I would like to thank my main PhD mentee J\o rn Scharlemann for taking me up as a PhD student and guidance along the way. 
I hold especially

Especially the dinners I hold our pub hikes, dear in memory
I very much enjoyed and our hiking trips around the South Downs
To this day I still wonder how and when you got past me at the Beachyhead Marathon and I wellcome a  
Thanks for being sometimes patient with me and showing that attention to detail in writing which I am still learning.

I think this is something I will take with me
I will always be proud to be Scharlemann lab PhD graduate \texttt{#2}.

Pedram
I always
I remember the words y
I hold the words you gave me the start of my PhD dear, that the success of a PhD is not measured by accomplishments, but by learning how to think.
I hope I have succeced learnt to think a bit more.
And hope greeting to your kids and Lord Rusby

Next, Lab 
Dan
Claudia
Owen

Rich
Edwin

Mijke, Jim And Josie

Other member of the deparment
Beth N
Gigi Toby
Veronica,

Mammal crew

Adam, Alan, Chris?

Rasmus (exchange)
Jury (exchange)
Jenny
Beth G
Paul D
Tom
Tanya and 
Will Gilks

and all other members of the department.

I would also like to acknowledge and thank my family in supporting me and assisting me to get where I am today. Danke Mutsch f\u er deine welt und ordnung zu hinterfragen sowie
mutsch, vatsch, brutsch
ganze familie

Finally, I would like to thank the kitchen personnel of the Swann Inn, Falmer for providing me that steady influx of declicious british cousine throughout my PhD. 